### 2023-09-08 paper1_figures.Rmd
sabina's version 08.09.23

---
title: "paper1_figures"
output: html_document
date: "2023-09-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

FIGURE 1: diagram of papers' selection process
``` {r}
library(DiagrammeR)
library(webshot)
DiagrammeR::grViz("digraph {
graph [layout = dot, rankdir = TB]

node [shape = rectangle]        
rec1 [label = 'Papers identified from search engines
n = 5210']
rec2 [label = 'Duplicates
n = 624']
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586']
rec4 [label = 'Papers not suitable
n = 4321']
rec5 [label = 'Papers read completely
n = 265']
rec6 [label = 'Papers not suitable
n = YYY']
rec7 [label = 'Studies included
n = XXX']

# edge definitions with the node IDs
rec1 -> rec2
rec3 -> rec4
rec5 -> rec6
rec1 -> rec3 -> rec5 -> rec7
}",
height = 500)
```