### 2023-10-05 paper1_figures.Rmd
final numbers 04.10

---
title: "paper1_figures"
output: html_document
date: "2023-09-06"
---

FIGURE 1: diagram of papers' selection process
``` {r}
library(DiagrammeR)
library(webshot)
DiagrammeR::grViz("digraph {
graph [layout = dot, rankdir = TB]

node [shape = rectangle]        
rec1 [label = 'Papers identified from search engines
n = 5210']
rec2 [label = 'Duplicates
n = 624']
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586']
rec4 [label = 'Papers not suitable
n = 4333']
rec5 [label = 'Papers read completely
n = 253']
rec6 [label = 'Papers not suitable
n = 157']
rec7 [label = 'Studies included
n = 96']

# edge definitions with the node IDs
rec1 -> rec2
rec3 -> rec4
rec5 -> rec6
rec1 -> rec3 -> rec5 -> rec7
}",
height = 500)
```