### 2023-10-10 paper1_figures.Rmd
10.10

---
title: "paper1_figures"
output: html_document
date: "2023-09-06"
---

FIGURE 1: diagram of papers' selection process
``` {r}
library(DiagrammeR)
library(webshot)
DiagrammeR::grViz("digraph {
graph [layout = neato, rankdir = TB]

node [shape = rectangle]        
rec1 [label = 'Papers identified from search engines
n = 5210', pos='0,0!']
rec2 [label = 'Duplicates
n = 624', pos='3,0!']
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586', pos='0,-1!']
rec4 [label = 'Papers not suitable
n = 4333', pos='3,-1!']
rec5 [label = 'Papers read completely
n = 253', pos='0,-2!']
rec6 [label = 'Papers not suitable
n = 157', pos='3,-2!']
rec7 [label = 'Studies included
n = 96', pos='0,-3!']

# edge definitions with the node IDs
rec1 -> rec2
rec3 -> rec4
rec5 -> rec6
rec1 -> rec3 -> rec5 -> rec7
}",
height = 500)
```

``` {r}
library(DiagrammeR)
library(webshot)
DiagrammeR::grViz("digraph {
  graph [layout = fdp, rankdir = LR]

  node [shape = rectangle]
  rec1 [label = 'social identity', pos='0,0!']
  rec2 [label = 'cooperation', pos='4,0!']
  rec3 [label = 'social norms']
  rec4 [label = 'communication']
  rec5 [label = 'trust']

  rec1 -> rec2
  rec1 -> rec3 -> rec2
  rec4 -> rec3
  rec4 -> rec1
  rec1 -> rec5 -> rec2
}", height = 500)
```