### 2023-10-11 paper1_figures.Rmd
11.10

---
title: "paper1_figures"
output: html_document
date: "2023-09-06"
---

FIGURE 1: diagram of papers' selection process
``` {r}
library(DiagrammeR)
library(webshot)
DiagrammeR::grViz("digraph {
graph [layout = neato, rankdir = TB]

node [shape = rectangle]       
rec1 [label = 'Papers identified from search engines
n = 5210', pos='0,0!', width = 4, height = 0.5]
rec2 [label = 'Duplicates
n = 624', pos='4,0!', width = 2, height = 0.5]
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586', pos='0,-1!', width = 4, height = 0.5]
rec4 [label = 'Papers not suitable
n = 4333', pos='4,-1!', width = 2, height = 0.5]
rec5 [label = 'Papers read completely
n = 253', pos='0,-2!', width = 4, height = 0.5]
rec6 [label = 'Papers not suitable
n = 157', pos='4,-2!', width = 2, height = 0.5]
rec7 [label = 'Studies included
n = 90', pos='0,-3!', width = 4, height = 0.5]

# edge definitions with the node IDs
rec1 -> rec2
rec3 -> rec4
rec5 -> rec6
rec1 -> rec3 -> rec5 -> rec7
}",
height = 500)
```

``` {r}
library(DiagrammeR)
library(webshot)
DiagrammeR::grViz("digraph {
  graph [layout = fdp, rankdir = LR]

  node [shape = rectangle]
  rec1 [label = 'social identity', pos='0,0!']
  rec2 [label = 'cooperation', pos='4,0!']
  rec3 [label = 'social norms']
  rec4 [label = 'communication']
  rec5 [label = 'trust']

  rec1 -> rec2
  rec1 -> rec3 -> rec2
  rec4 -> rec3
  rec4 -> rec1
  rec1 -> rec5 -> rec2
}", height = 500)
```