### 2023-10-12 paper1_markdown.Rmd
12.10

---
title             : "Improving the understanding of cooperation in commons through social dilemmas' literature: A scoping review"
shorttitle        : "Improving the understanding of cooperation in commons"

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "University of Bern"

authornote: |

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  According to PRISMA-ScR: Provide a structured summary that includes (as applicable) background, objectives, eligibility criteria, sources of evidence, charting methods, results, and conclusions that relate to the review questions and objectives

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "cooperation, social dilemmas, commons"
wordcount         : ""

bibliography      : CommonPaths.json

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
always_allow_html : true
#bibliography: references.bib
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

During last years, the interest in alternative ways of social organization augmented drastically because of the social and ecological crises we are facing [@Euler2018]. Commons are one of this possible solutions, since they represent an alternative way of organizing resources who does not involve the classic private and public model [@Ostrom1990]. In her seminal work, Ostrom refers to commons (or "common-pool resource") as a resource system equally shared between a self-managed community of users, also because of the difficulty of excluding somebody from benefiting from its use [@Ostrom1990]. Beyond the material aspect of the commons, the notion of social practices (referred as commoning) is also essential [@Euler2018]. Commoning is the process through which people manage to find solutions for their needs by self-organizing independently from the state or the market. Community is therefore engaged in defining rules to successfully manage a shared resource [@Meroni2022], and that is why taking into account the specific community who is commoning is also essential to fully understand a commons [@Dove2019]. @Bollier2019 are even more radical: for them, the term "commons" should not be defined as a resource at all, but as "living social systems" who allow people to self-organize to address shared problems. Finally, "commons can [...] be understood as collaborative arrangements for value production processes that are based on participatory principles" [@Seravalli2018a, p. 1].

Commons are currently receiving more and more attention because they function outside of the capitalist system, they respect the Earth by giving mutual benefits to their members [@Meroni2022], and they could be the needed change toward more sustainable and fair economy and society [@Seravalli2018a]. In fact, having small groups with a high perception of risk and a minimal uncertainty in collective goal, possibly focused on regional issues, could significantly improve the probability of success in attaining common goals related to sustainability [@Pacheco2014]. Moreover, it is easier to coordinate on a local, rather than global, level [@Carpenter2011]. Broadly, commons represent a possible solution simply by showing that alternatives to the classic model are possible [@Baibarac-Duignan2023].

However, in order for the commons to be successful its members have to cooperate, i.e to incur a personal sacrifice for the greater good [@Kollock1998], but this cooperation can be difficult to achieve [@Carpenter2011]. This is explained by two characteristics of the commons' social dynamics. First, there is a strong asymmetry, in the sense that the collectivity has a strong impact on the individual, but a person's decisions have a smaller effect on the collective [@Kramer1984]. Therefore, if people are uncertain about others' cooperation, the best solution is pursuing their personal self-interest. Secondly, in a commons situation, every member faces a conflict between individual, short-term interests and collective, long-term ones [@Hardin1968]. For instance, a fisher could be tempted to fish in large quantity from a common basin, but if every other fisher did the same there would not be enough fish for the following seasons. According to @Hardin1968 and his "Tragedy of the Commons", individuals' greed to maximize their personal payoff will inevitably lead to the destruction of the common resource. For these reasons, it is essential to understand why people do (not) cooperate in commons, and which factors can impact cooperation.

## The study of cooperation

It exists different approaches to the study of cooperation. For instance, classical economic theories consider individuals as completely rational and always maximizing (expected) utility [@Edwards1954]. Thus, according to these theories, individuals are not supposed to behave altruistically, namely to incur a personal cost while benefiting others [@Wittek2015]. Another theoretical approach toward cooperation is the evolutionary perspective, who suggests that individuals can show cooperation, but only when doing so bring them a personal advantage. For example, the theory of competitive altruism says that acting for the others' good could be a signal of a person's pro-sociality, and therefore improve this person's reputation [@Roberts1998]. Moreover, the capacity to endorse a personal cost communicates strength and wealth, socially valued characteristics [@Zahavi1995].

Cooperation has also extensively been studied through social dilemma. In such situations, each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. However, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980]. This approach is very useful to gain insights into how cooperation works because it provides a standard experimental paradigm to test under which conditions people cooperate [@Balliet2011]. While economic and evolutionary theories say that individuals are always pursuing their self-interest (either by defecting to maximize their personal payoff, either by cooperating to increase their social status and reputation), social dilemma experiments' results show that it is not always the case. First, it is observed that a significant amount of people are "strong reciprocators": they are willing to use their personal resources to reward cooperators and punish defectors, even if they are not certain that their altruistic behaviour will be rewarded [@Fehr2002; @Gachter2009]. Moreover, according to the Greed Efficiency Fairness hypothesis, people's greed to pursuit their personal interest is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. These results shows that cooperation is not always driven by personal interest and that people care about others' and environment's welfare [@Torsvik2011; @VanVugt2009].

In addiction to that, there are evidences suggesting that people do not always thoroughly analyze the context to take a rational decision. In fact, cooperation seems to be the intuitive option in social dilemma experiments, while defection needs more thinking [@Kieslich2014]. Moreover, people can also cooperate in part because they are confused about the situation [@Andreoni1995]. In fact, @Fosgaard2017 showed that, in public good games, there is a considerable amount of participants that do not understand correctly the strategies to maximize the personal or the collective payoff, and these misconceptions can vary depending on how the game is presented. Moreover, there are individual differences in the level of rationality, i.e. the ability to think analytically at the situation [@Lang2018]: individuals scoring low in rationality cooperate more than individuals with an high score. Therefore, people are not always rational in the economic, individualistic way [@Kollock1998], and there are multiple factors that can influence their behaviour in social dilemmas situations [for a review, see @VanLange2013].

## Current review

Like social dilemma experiments challenge economic and evolutionary assumptions about cooperation, so do observations of real-life commons with Hardin's "Tragedy of the Commons": in fact, there are many examples of successful commons [e.g. @Ostrom1990; @Franzen2019]. However, cooperation in commons has not been systematically studied like it has been in social dilemmas, and there is therefore a lack of knowledge about the factors that could enhance cooperation in this context. A way to start to fill this gap is to apply results of social dilemma to the commons. Linking this two topics is pertinent for at least three reasons. First, social dilemmas and commons share the same basic conflict between personal and collective interests [@Hardin1968; @Dawes1980]. For this reason, @Sussman2016 used the term of "commons dilemma" and considered commons to be a particular case of social dilemma. Secondly, cooperation and defection in social dilemma laboratory experiments can predict people's behaviour in a commons situation [e.g. @Reindl2019; @Ros-Galvez2015]. Finally, others studies already used the social dilemma framework to study real-life commons [e.g. @Hartl2022].

Based on these evidences, the aim of the present review is to use social dilemma literature to gain insights into cooperation in commons by (a) identifying individual factors that can shape it and (b) understanding which techniques groups can implement to encourage their members to cooperate. It is important to note that the point (b) will focus exclusively on bottom-up strategies that groups can adopt, and not on top-down solutions like privatization or centralized control of the resource. In fact, it well demonstrated that members of a commons can efficiently manage it without the intervention of an external institution [@Larsson2016]. Moreover, bottom-up solutions are often more efficient than the two others [@Ostrom1990].

# Methods

Given the exploratory nature of this paper, a scoping review approach is used [@Munn2018]. To conduct it, different scientific publications were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust", and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore journal articles, conference papers and book chapter could be included. This allows us to have an idea as complete as possible about the existing literature on cooperation in social dilemmas and commons. Moreover, there were not any chronological or geographical limitations, once again to provide a broader vision of the topic. There were however some selection criteria concerning the papers' methodology and content. First, as mentioned in the introduction, papers investigating how an external entity/institution shapes cooperation were not included. Secondly, human participants were needed, so papers who studied social dilemmas with an agent-based model methodology were also excluded. Moreover, in case of experimental or correlative studies, only cases with adult participants were selected.

Initially, 5210 papers emerged from the key phrases search. After deleting the duplicates, the count was of 4586. Based on titles and abstracts, 4321 articles were eliminated, leaving 253 texts to examine. After a full-text evaluation of the remaining papers, 90 were selected for the present review. Figure 1 resumes this selection procedure.

```{r}
library(DiagrammeR)
library(webshot2)
DiagrammeR::grViz("digraph {
graph [layout = neato, rankdir = TB]

node [shape = rectangle]       
rec1 [label = 'Papers identified from search engines
n = 5210', pos='0,0!', width = 4, height = 0.5]
rec2 [label = 'Duplicates
n = 624', pos='4,0!', width = 2, height = 0.5]
rec3 [label = 'Papers screened based on titles and abstracts
n = 4586', pos='0,-1!', width = 4, height = 0.5]
rec4 [label = 'Papers not suitable
n = 4333', pos='4,-1!', width = 2, height = 0.5]
rec5 [label = 'Papers read completely
n = 253', pos='0,-2!', width = 4, height = 0.5]
rec6 [label = 'Papers not suitable
n = 163', pos='4,-2!', width = 2, height = 0.5]
rec7 [label = 'Papers included
n = 90', pos='0,-3!', width = 4, height = 0.5]

# edge definitions with the node IDs
rec1 -> rec2
rec3 -> rec4
rec5 -> rec6
rec1 -> rec3 -> rec5 -> rec7
}",
height = 500)
```

## Papers' description

The 90 included papers were published between 1984 and 2023, with a median of 2015. Almost all of the selected texts were journal articles, with the only exception being a book chapter. Most of them were experimental studies (*N*=61), followed by literature review and meta-analysis (*N*=14) and correlational studies[^1] (*N*=14). The remaining paper was a qualitative study.

[^1]: Studies were at least a variable is manipulated by researchers are part of the category of "experimental studies", while "correlational studies" refers to papers using quantitative analysis, but without manipulations.

The papers' topics can be divided in 4 categories: first, "social dilemma setting" (*N*=76) refers to high controlled laboratory experiments where participants had to take cooperative decision. Usually some variables were manipulated to investigate the influence of a specific factor on cooperation [e.g. @Zhang2019]. The type of paradigm used was most of the time a public good game (where participants could decide how much to contribute to a public good, e.g. @Torsvik2011) or a common resource game (where participants could decide how much to take from a common resource, e.g. @Nockur2020). Moreover, the type of game could vary during the experiment [e.g. @Fosgaard2019]. The second category is "real-life social dilemma situation" (*N*=7), and investigated cooperation in different real-life situations with a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Thirdly, "real-life commons" (*N*=4) refers to papers who studied cooperation with participants that were members of a commons in their everyday life. Finally, papers in the category "commons modeled by a social dilemma experiment" (*N*=3) used a laboratory social dilemma paradigm, but with clear references to a concrete common. For instance, in @Mosler1993, participants had to imagine to be fishermen who shared a common basin.

Most of the papers concerned psychology and economy. In fact, thirty-five papers were published in psychology journals, such as *Group Processes & Integroup Relations* or *Psychological Bulletin*, and fourteen in economic journals (e.g. *Experimental Economics*). Additionally, some papers (*N*=8) were published in interdisciplinary journals such as *Journal of Economic Psychology.*

Finally, most of the studies were conducted in Europe (*N*=39) (especially in Germany (*N*=12) and in the Netherlands (*N*=9)), followed by North America (*N*=20, *N*=19 of which in the USA) and Asia (*N*=9, *N*=6 of which in China). Moreover, *N*=6 studies were set in more than one country, and *N*=14 were meta-analysis or literature reviews. The two remaining studies took place in Australia and in Colombia, respectively.

# Individual factors

## Gender

The effect of gender on cooperation is controversial, and studies that investigated it provided contradictory results: sometimes men cooperated more [@Dorrough2019], sometimes women did [@Peshkovskaya2017], and sometimes there was no difference between genders. In general, meta-analysis found a non significant effect size for this difference [e.g. @Balliet2011; @Spadaro2023]. This indicates that gender differences in cooperation are very flexible and can vary depending on the context [@Spadaro2023]. Therefore, when a significant effect is observed, it means that gender has somewhat been made salient by the situation [@Sell2009].

One of the most observed moderator of this effect is the interaction partners. In fact, in a social dilemma experiment, women were more cooperative toward strangers than men, while men cooperated more in a group of friends. Moreover, after a socialization phase where strangers could interact, men cooperated more than women [@Peshkovskaya2017]. These results seem to indicate that men accord more importance at the relationship with other members than women. @Gomez-Ruiz2020 also got to the same conclusion by showing that the creation of a team identity enhanced men's cooperation, but had no effects on women. Moreover, in a social dilemma with many iterations, men, but not women, tended to increase their cooperation over time [@Balliet2011].

The partners' gender also have an impact on women's and men's cooperation. In general, groups composed only by men are more cooperative than only-women groups. This is explained by the fact that, in general, women are less achievement oriented and more socially oriented than men: they are therefore less concerned by the maximization of their personal payoff and more about relative payoff. Thus, the defection strategy allows them to play defensively and to avoid earning less than the other players [@Colman2018]. Moreover, men displayed higher levels of gratitude and trust in same-sex groups than women [@Peshkovskaya2019]. On the other side, women cooperate more in mixed group. In fact, when both men and women are present, gender stereotypes are activated and people tend to adapt their behaviour to them. This leads women to behave in a more prosocial, altruistic way and men to be more self-oriented and competitive [@Balliet2011]. The same result is also found by @Peshkovskaya2019, who addiotionally demonstrated that women reported a stronger social connection with their partners in mixed than in same-sex groups. The overall group composition also have an impact on cooperation: in general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023]. However, @Barrero-Amortegui2021 showed that the level of resource exploitation is linked to the number of women in a group, but that this link is mediated by the signature of an agreement about extraction. More precisely, in their experiment, groups with a majority of women extracted more resources from a common good (and thus cooperated less) before the signature of the agreement, but also showed the largest reduction of the exploitation level after the agreement was signed. Therefore, even the effect of group composition seems to be subjected to moderator variables.

Another reason why gender differences in cooperation are often very small is that fear and greed motives are confused in a classical social dilemma situation [@Simpson2003]. In this contexts, a feeling of fear arises when one's cooperation could be exploited by others' free-riding, while greed is the feeling associated with the temptation to free-ride on others' cooperation. According to social psychology and evolutionary theories, the author expected women to defect more when fear incentives were present, and men to defect more with greed incentives. Although he could corroborate only the second hypothesis, his intuition about women being more sensitive to the fear of being exploited is confirmed by other studies. For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

Finally, some of the analyzed papers presented interesting secondary results. First, against their authors' expectations, some intercultural studies showed that the gender difference in cooperation is not explained by the level of gender equality of the country [@Dorrough2019; @Spadaro2023]. Secondly, in an experiment not involving a social dilemma, women were significantly more altruistic than men [@Dorrough2019]. According to the authors, this underlines the importance of distinguishing between prosociality and strategic interactions that characterize social dilemmas. Therefore, in a situation where classical attributes of social dilemmas are not explicit, gender difference in cooperation could not be subjected to the same moderator effects described above. Lastly, according to @Sell2009, it is not possible to think about gender without relying it to social status, with men representing the high-status group. Therefore, gender differences in cooperation could also be caused by social status: to have a better understanding of this hypothesis, it would be interesting to investigate the difference in cooperation between two other social groups with a different status (e.g. high and low socioeconomic status).

In conclusion, "at times women are more cooperative than men, less cooperative than men, or equally as cooperative as men. These variations are explained, in large part, by structural differences and identities triggered by those differences" [@Sell2009]. The context in which men and women interact is therefore essential to predict if a gender will cooperate more than the other, and which one.

## Social status

Few studies have investigate how social status impact cooperation, and their results do not allow to draw a clear conclusion about this effect.

For instance, advantaged members (i.e. high-status class) tended to cooperate less than disadvantaged ones [@Aksoy2019]. In fact, when participants did not receive the same endowment at the beginning of the experiment, contributions of "rich" and "poor" members converged [@Kingsley2016]. In other words, proportionally to their resources, low-status participants cooperated more. However, in their review, @VanLange2013 identified some studies that observed the opposite effect, namely that wealthier people (or those who benefited more from a common resource) contributed more to the common good than the other.

Social status can also create inter-class conflicts. In fact, people cooperated more with the members of the same social class than with other classes' members [@Aksoy2019]. Moreover, a form of solidarity is observed between low-status members: in a study where participants had the opportunity to punish others for a non-cooperative behaviour, low-status members punished significantly less other low-status participants [@Chen2022]. Also, in an experiment where participants were randomly advantaged during the interaction, @Camera2020 found that subjects who received a lot of advantages were discriminated by others, and especially by the ones who were rarely advantaged.

Regarding the overall group composition, results are not consensual. On one hand, @Aksoy2019 found that groups with an inequality of status between members displayed lower levels of cooperation than homogeneous groups. On the other hand, economic and socio-cultural heterogeneity, although being negatively correlated with in-group trust, does not have a direct effect on a common-pool resource well-being [@VanKlingeren2021].

In conclusion, social status impacts cooperation, but it seems to have multiple effects. Moreover, as mentioned in the previous section, @Sell2009 affirmed that, when investigating gender differences in cooperation, it is impossible to disentangle the effect of social status. Similarly, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that therefore it is difficult to separate their effects on cooperation.

## Group identification

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group [e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020]. For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Group identification can also esaily be implemented in a laboratory experiment, and it has still a strong positive effect on cooperation [@Wit1992]. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

However, the fact that not all members identify with their group to the same extent can lead to inequalities on cooperation. For example, @VanVugt2009 mentioned that high identified individual can even compensate, by cooperating more, for other in-group members' defection. The same result was also found by @Arora2016: high identified individuals cooperated more than low identified ones when faced to an in-group participant who defected (and this difference was not found when the defector was not a group member). However, this compensation came with a cost, since participants who compensated for an in-group defector reported lower trust toward this person.

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption, while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

However, despite the observed beneficial effects of inducing an intergroup competition, in a real-life situation this could be harmful. This is especially true in case of resources that are shared between several groups, which are at greater risk of over-exploitation in case of intergroup competition [@VanVugt2009]. A possible solution to maintain the positive effects of group identification, while eliminating the risk of over-exploitation caused by the competitive context, is to prime a superordinate identity [@VanVugt2009]. In fact, intergroup cooperation can be enhanced by making salient the common characteristics of the different groups, instead of their differences [@Kramer1984]. Another effective way to create a superordinate identity is to induce the feeling of a common fate, for example through a collective risk. In fact, a common fate allows a group to become real [@Zhang2019], and can therefore erase intergroup competition.

Finally, social identity does not only have a main effect on people's cooperation, but can also moderate other factors' impact. For example, monetary incentive are more effective at enhancing cooperation of low identified individual than of high identified ones [@VanVugt2009]. Moreover, a previous group failure can decrease cooperation between low identified member, but this effect is not observed in people who have a strong identification with their group [@Jackson2011; @Jackson2012].

In general, social identity seems to have a positive and largely observed effect on individuals' cooperation. This is due to the fact that humans have a very developed sense of belonging to social groups, and that this has an impact on their willingness to help other members of their group [@VanVugt2009]. Since priming a common social identity is quite simple (e.g. @Kramer1984; @Zhang2019), this could be a very powerful technique to enhance cooperation in an easy and cheap way.

## Values and personality traits

Personal values and personality traits can play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation is very consensual: in a meta-analysis, 81 out of the 82 papers examined reported than pro-social cooperated significantly more than pro-self [@Balliet2009].

However, this general effect is subjected to several moderators. For example, pro-self individuals cooperated more in a game where economic incentives to cooperation were present than in a game without such incentives, while pro-social did not displayed this difference and showed high levels of cooperation in both contexts [@Emonds2011]. A similar effect is also described by @Balliet2009, who found a stronger effect of SVO in unpaid games than in paid ones, because, when monetary rewards were present, pro-self increased their cooperation. Therefore, pro-self individuals also cooperate, but only when cooperation serves their personal interest. In fact, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998].

Moreover, SVO has a stronger effect if there is an uncertainty about the number of individuals having access to a common resource: when the number of other participants was not known, pro-social, but not pro-self, reduced their harvesting comparing to a group-size certainty condition [@DeKwaadsteniet2008]. The same authors demonstrated that this result is explained by the fact that pro-social individuals want to maximize the joint outcome, while pro-self only think at their own personal interest. Also, SVO effect is partly explained by trust, i.e. the extent to which people think that their interaction partners will also cooperate. In fact, in @VanLange1998, pro-social individuals with a low level of trust did not make different decisions than pro-self individuals, while the most cooperative were pro-social with an high level of trust. Finally, @Lu2019 showed that the intuitive choice of pro-social individuals is cooperation, while for pro-self is defection. However, pro-self can cooperate more if they have the possibility to deliberate about the situation, because they can realize that the best strategy to maximize the joint (and thus personal) outcome is cooperation.

Other studies also focused on how much people are self- or social-oriented by using other constructs. For example, @Liu2009a demonstrated that priming a common fate to induce a context of interdependence significantly increased the cooperation level of participants with an independent self, but had no effect on participants with an interdependent self (who already displayed high levels of cooperation). Moreover, in a social dilemma experiment, competitive, individualistic people were motivated just by personal gain, while people who saw themselves as a part of a group were more interested on maximizing the income of each member, even at the cost to have a smaller personal payoff [@Probst1999]. Also, high allocentric (i.e collectivist) individuals cooperated more than low allocentric ones when the perceived social norms are cooperative: this means that they are more sensitive to social cues and therefore they adapt their behaviour to social norms [@Chen2007].

Finally, there are other personal values and traits that can affect cooperation. For example, pro-environmental values are negatively correlated with harvesting from a common resource, but only when the resource is scarce; therefore these values are more likely to motivate difficult, rather than easy, ecological behaviours [@Sussman2016]. Levels of harvesting from a common resource are also lower between left-wingers than right-wingers [@Fosgaard2019]. Moreover, humility and honesty are significant predictor of cooperation [@Kieslich2014], and this is due to the fact that people with such characteristics manage to resist to the temptation to exploit others' contributions [@Hilbig2018]. Self-control also plays a role, in the sense that more self-control leads to a greater cooperation [@Kocher2017]. Lastly, people with an higher justice sensitivity make bigger contributions to a public good [@Schlosser2018].

In conclusion, values and personality traits have a large impact on human cooperation, and it is thus very important to take them into account when investigating this phenomenon. Interestingly, some social psychological techniques (e.g. elaboration) allow to modify an individual's values [@Sussman2016]. Such techniques could therefore be used to improve cooperation.

## Trust and beliefs about others' contribution

In social dilemma situations, trust can be defined as the "expectations of others' benevolent motives in situations that involve a conflict between self and collective interests" [@Balliet2013, p.1]. Trust can be dispositional or a state; the former is an individual's general trait, and is usually measured with personality scales, the latter is linked to the context, and can be operazionalized by expectations about a partner's behaviour in a specific situation [@Balliet2013]. The two are positevely correlated [@Lubke2021], and therefore both have a positive general effect on cooperation [@VanLange2013; @Balliet2013]. In fact, high levels of trust are associated with more contributions to a public good [@Kocher2015], and a participant's contribution in a public good game can be predicted by her beliefs about others' contributions [@Smith2013].

Several studied confirmed this relationship. For instance, @Chaudhuri2002 demonstrated that people who cooperated more in a first public good game trusted more their partner in a following interaction than people whose main strategy was defecting. Moreover, people that were pessimistic about others' contribution cooperated significantly less than the optimistic [@Oyediran2018], and those who free-rode in a public good game, compared to cooperators, believed that others contributed less [@Kocher2015]. Also, people were more willing to purchase actions for a CO2 mitigation initiative when they believed that the majority of others had also bought them [@Sturm2019]. Finally, in a study about self-administered outlets in Swiss farms (i.e. outlets where nobody monitors if people pay or not their purchase), @Franzen2019 found that a farmer's general trust in other people is a significant predictor of the probability of having such an outlet. Concerning the moderators, trust matters more in situations involving a large, rather than small, conflict between personal and collective interests [@Balliet2013], and when people are uncertain about others' intentions [@VanLange2013].

People can also display different levels of trust in the "reliability, effectiveness and legitimacy of public institutions" [@Rompf2017, p. 2]. This authors showed that institutional trust is associated with the probability of recycling, because knowing that sanctions can be efficiently applied can motivate people to adapt to norms. Moreover, the effects of perceived costs and benefits of the action are less predictive in people with high institutional trust.

Finally, trusting others (and the institutions) is essential to display higher level of cooperation. However, general beliefs about others' cooperation are quite low compared to the actual behaviours, and this trust gap can undermine cooperation, since people's actions are significantly predicted by their beliefs [@Lubke2021]. In order to increase trust and, consequentely, cooperation, it is therefore it is essential to provide people with reliable information about others' behaviour.

# Group factors

## Incentives

Cooperation rates can be modified by introducing incentives, who can be either punishments or rewards[^2]. In the litterature, punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation is very consensual. In general, when there is the possibility for participants to punish others members based on their contributions, cooperation is higher than when this is not possible [e.g. @Pfattheicher2018; @Balliet2011a]. Moreover, the presence of a punishment system increases the perception of fairness [@Nockur2021].

[^2]: Since the present paper's interest is to investigate how the internal organization of a group can affect cooperation, only papers that studied situations where the participants could give rewards or punishments to each-others are included. Incentives given by third party have also an effect on cooperation, but they are not included in this review.

However, this main results is subjected to several moderator factors. First, the punishment have a bigger impact on cooperation if its implementation is costly for the punishers than if it is free [@Balliet2011a], but at the same time the cost should not be too high [@Egas2008]. Moreover, punishments are more effective if the interaction is among friends than among strangers, and if people know that they will have to interact repeatedly with each other [@Balliet2011a]. In fact, in short interactions, if punishment is costly, individuals who punish others are the ones with the lowest personal income [@Milinski2012], and this could discourage the application of the fine. Also, an experiment showed that a punishment system increased cooperation in groups where all members had the same amount of resources, but not in groups where income was heterogeneous [@Kingsley2016]. In fact, @Chen2022 found that, in groups where the initial endowment was heterogeneous, participants that received less didn't punish the others "low-status" members. This solidarity could therefore encourage free-riding. Moreover, the effect of punishment is stronger when only cooperators have the power to punish defectors, than when everybody can punish [@Xiao2016]. Finally, a plain situation is also needed to have a stronger effect of punishment: the link between behaviour and outcome must be clear, without random factors affecting the final outcome [@Xiao2016], and the punishment must be highly effective [@Egas2008].

However, punishment must be applied with carefulness: @Rockenbach2019 compared the punishment strategy of different groups, and found that the most successful ones (whose members had the highest payoff at the end of the experiment) punished with caution and not in early stages of the game. In these groups, a norm of when and how punishment should be applied seems to exist. Moreover, punishment has also two main negative sides. First, it could entails a second-order social dilemma [@Milinski2012], since cooperative participants can free-ride on the effort of other cooperators who punish non-cooperators. Secondly, anti-social punishment can arise. This term refer to the punishment of cooperators by non-cooperators, and it is present when all participants can punish independently of their contributions [@Xiao2016]. In this context, defectors can see cooperators as rule breakers, and anti-social punishment is a sort of retaliation [@VanDijk2015].

A possible solution could be to implement a system of democratic punishment, where a fine is given only if a majority of participants agree. In fact, compared to a peer punishment system where participants decide individually if and who punish, in a democratic system anti-social punishment is less present, general punishment level is lower, and final payoff and cooperation rates are higher [@Pfattheicher2018].

Rewarding systems, on the other side, are less investigated. Globally, they have a positive effect on cooperation [@Balliet2011a]. However, some studies found that they were more efficient than punishment [e.g. @Chen2022], while others came to the opposite conclusion [e.g. @Milinski2012] or found no difference between the two systems [e.g. @Balliet2011a].

In a meta-analysis, @Balliet2011a found that rewards had a stronger effect on cooperation if their implementation was costly rather than free, and in iterated, rather than single, interactions. Moreover, rewarding was more effective in situations where participants had to decide how much take from a common resource, compared to when they had to decide how much contribute to a common good.

In general, incentives seem to be a powerful tool to enhance cooperation, but they have a dark side. In fact, they reduce participants' intrinsic motivation [@VanDijk2015], by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. In fact, @Schlosser2018 demonstrated that people's level of justice sensitivity could predict their contributions to a public good game, but this link disappeared when a punishment system was introduced. Also, when incentives are present people think that others cooperate just because of them, and this undermines their trust in other members[@Irwin2014], and especially the trust in the fact that others are internally motivated to cooperate [@Mulder2006]. Therefore, if incentives are removed, cooperation drops [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. For example, the expectation of honour or shame (after cooperating or defecting, respectively) is enough to enhance cooperation [@Jacquet2011]. Moreover, expressing disapproval toward defectors can boost cooperation by inducing more compliance to the cooperative social norm [@Nelissen2013]. Even an ostracism system where participants can vote to exclude other members from the game is very effective in increasing cooperation [@Maier-Rigaud2010]. In general, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

To conclude, economic incentives can have a positive effect on cooperation, but their efficient implementation must be insured all along the interaction, otherwise cooperation rates would drop. On the other hand, social incentives seem to have the same positive effect, and their removal do not cause a reduction of cooperation. However, there is a lack of studies focusing on this kind of sanctions.

## Communication

Communication has a global positive effect on cooperation [e.g. @Balliet2010; @Bicchieri2002], but this effect is subjected to several moderators. First, talking is more powerful than sending written messages, because it allows non-verbal communication and facilitates the social norms formation process. Moreover, communication's effect is stronger in large groups: when many people are present cooperation rates are lower, so communication between members can buffer this negative effect [@Balliet2010]. Finally, if communicate with others requires a personal cost, the effect is weaker than if communication is free [@Janssen2014].

The content of the messages is also essential. @Koessler2021 identified four steps that need to be present when participants in a social dilemma situation communicate: problem awareness, identification of strategies, agreement and ratification. In another study, groups who used communication to foster a common social identity, promote cooperative social norms and ensure that all the players publicly committed displayed higher rates of cooperation than other groups [@Adams2022]. However, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Also, the increase of cooperation through communication can be explained by the normative content that people are shaping while exchanging messages [@Janssen2014], and by the fact that it allows participant to reach a consensus about the best strategy [@Hopthrow2005]. Moreover, even if communication also helps to create a social identity, the effect of social norms is a stronger mediator of the link between communication and cooperation [@Bicchieri2002].

Another factor that influences the effect that communication has on cooperation is whether individuals exchange messages before, during or after the cooperative interaction phase. Concerning the post-game communication, @Kumakawa2013 found that receiving a feedback about the cooperation choice made previously can increase cooperation in the following interactions, especially if the feedback is negative. However, just the fact of knowing that a discussion will follow is not enough to enhance cooperation, and can even have a counterproductive effect, because forcing people to think about a future discussion make them reflect more on the situation and thus act more rationally [@Torsvik2011]. On the other hand, communication before or during the interaction have both a positive impact on cooperation, and the difference between their effect is not significant [@Balliet2010].

Finally, several studies investigated the effect of gossip as a form of communication that can increase cooperation. Gossip is defined as a spread of reputational information about another person [@Feinberg2014], and it increases cooperation in two ways: first, if people know that gossip is possible, they tend to cooperate more to avoid the risk of being the target of such gossip and therefore of being excluded from following interactions. Secondly, having information about the past behaviour of potential cooperation partners can help to select the most cooperative people in an interaction situation [@Feinberg2014]. Moreover, the same authors found that pro-social individuals gossiped more than pro-self. This shows that gossip can have the pro-social function of protecting others from a potential exploitation [@Feinberg2012]. However, gossip has also some self-interest effects: for instance, observing another subject being exploited and having the opportunity to gossip about the exploiter significantly reduced negative affects in @Feinberg2012 's participants. Moreover, providing correct information while gossiping helps to build a positive reputation [@Giardini2021].

In conclusion, when people have the possibility to communicate before, during or after a cooperative interaction, the cooperation level increases. Even if the strength of this effect can vary depending on the situation, communication never seems to have a negative impact on cooperation.

## Social Norms

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour [@VonBorgstede2018]. @Thogersen2008 studied the effect of social norms on daily social dilemma situations such as environmental behaviours, and found that both injunctive (what others approve) and descriptive (what others do) are positively related to behaviours. Moreover, cooperation can be maintained via social norms, because people who do not respect them are susceptible of social sanctions. In fact, if social approval is possible, people cooperate more [@Rege2004].

There are several explanations of social norms' effect on cooperation. Firstly, according to the appropriateness framework, when taking a decision people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. Therefore they need to define the situation as a cooperative task and to strongly identify with the group [@VanLange2013]. This leads to a greater accessibility of group's norms and consequently (if the norms are cooperative) to more cooperation. Testing this assumption, @Arora2012 showed that a greater identification with social group lead to higher levels of cooperation. The second explanation concerns conformity: in general people adapt their cooperation to others' behaviour, but this is usually explained by the reciprocity principle (cooperate enhances the probability that others will do the same, and therefore that the final payoff will be higher) [@Bardsley2005a]. However, these authors demonstrated that people conform to others' behaviour even if this has no impact on their personal payoff. Therefore, people have a general tendency to imitate others to follow social norms.

Finally, social norms can be modified by participants. For instance, @Weber2008 demonstrated that in several social dilemma experiment there are consistent cooperators, i.e. people who always cooperate. The presence of these people can increase other members' cooperation, and this effect is mediated by the perception of the social norms. Moreover, social norms can also be shaped via communication [@Bicchieri2002].

In conclusion, social norms have a significant effect on people's cooperation in social dilemma situations. Since they are quite easily modified [@Weber2008], shaping cooperative norms can be a powerful instrument to improve cooperation.

## Anonymity

There are not many studies on the influence of anonymity on cooperation, but, according to @Wang2017, most of the factors that can increase cooperation rely on a reduction of anonymity. For instance, it is not possible to punish defectors without knowing who they are. Therefore, in general it seems that non-anonymous situations can significantly increase cooperation and reduce defection [@Wang2017]. This can be explained by the fact that, when their actions are visible, people worry about others' approval [@Rege2004] and that, in non-anonymous situations, contributions can be a self-presentation strategy, because being perceived as pro-social can increase one's social status [@VanVugt2010a]. There are therefore more social incentives to cooperate in non-anonymous, rather than anonymous, situations.

Moreover, this effect is displayed especially in situations when individuals face a cost when cooperating and know that they will not get any personal benefit. For instance, @Butz2021 demonstrated that, in a public good game where a part of contributions was donated to a charity association, disclosing identity increased cooperation, but only if the donation was deducted from participants' payoff. In the same direction, @VanVugt2010a found that, when actions were public, contributions were higher if the public good was already attained or unattainable, and therefore if participants did not have a personal interest in cooperating.

Reducing anonymity in cooperative interactions can be very effective to encourage people to contribute more. A possible way to do this is to reduce group's size. In fact, according to @Romano2016, a larger group leads to de-individuation, because it is more difficult to understand who is defecting. Therefore, it is easier for members to feel anonymous and to perceive the social punishments associated with defection as less probable. Moreover, communication is easier in smaller groups [@VonBorgstede2018]. This assumption is confirmed experimentally by @Jiang2021, who showed than in smaller groups people contributed more often to a public good. Increasing cooperation through the reduction of anonymity is also possible by giving public feedbacks about people's cooperation [@Przepiorka2020].

Another effective method to reduce anonymity is to make people publicly promise to cooperate before the cooperative interaction. For instance, in a laboratory experiment modeling a common pool resource, people who publicly committed exploited significantly less the resource than people who committed in private, even if after committing they were not restricted in their harvesting possibilities [@Mosler1993]. This results was also confirmed by @Przepiorka2020, who showed that non-binding promises increased cooperation. Moreover, even a compulsory commitment has a beneficial effect on cooperation, even if the most efficient groups are those where all participants signed voluntary the pledge [@Koessler2021a].

In conclusion, reducing anonymity during interactions is essential to improve cooperation, because this can facilitate the implementation of others cooperation-enhancing techniques [@Wang2017]. Moreover, it exists several low-cost method to make situations non-anonymous.

# Discussion

Through a scoping review approach, the current study identified several factors that could impact cooperation in social dilemmas and in commons. Some of this elements can be related to Ostrom's principles to successfully manage a commons [@Ostrom1990]. In fact, her first statement says that commons need clearly defined boundaries: therefore, it must be clear who has the right to benefit from the common resource and who does not. This could be linked to the social identity construct, since only when an in- and an out-group exist social identity is salient and therefore an identification is possible [@Tajfel1974]. Even if Ostrom developed this principle on a more organizational level, the secondary effect of creating a social identity for commons' members can also contribute to improve cooperation. Moreover, individuals who are part of the group, and who are therefore affected by the commons' rules, must be able to participate in modifying them (principle 3). This statement shows the importance of social norms, and especially of injunctive ones, namely what is morally approved in a social group [@Cialdini1990]. Moreover, in order to modify such norms, people have to communicate [@Bicchieri2002]. Communication is also essential for principle 6, who posits that members of a commons must have access to low-cost arenas to resolve conflicts. In fact, communication helps to reach an agreement to the strategy that needs to be followed [@Hopthrow2005]. Finally, Ostrom's principles 4 and 5 underline the importance of incentives, more precisely in the form of a monitoring system implemented by members them-selves and a graduate sanctions system toward violators of the rules. The monitoring system refers to social incentives, and is linked to anonymity, since if people know that their behaviour can be observed, they are more willing to conform to social norms because they are worried about others' approval [@Nelissen2013; @Rege2004]. On the other hand, a sanction system can also have a positive impact on cooperation [@Balliet2011a], especially if it is graduate, so if sanctions are not too high at the first violation [@Rockenbach2019].[^3]

[^3]: Principles 2, 7 and 8 refer to elements that are not pertinent to the current review, and therefore do not match any of the factors identified.

Ostrom's assumptions are therefore confirmed, especially by factors at the organizational level. This is understandable when considering her economic approach to the study of the commons. However, taking into account the social dynamics and the individuals who interact within the commons is also essential to have a better understanding of how these communities work [@Euler2018; @Nightingale2011]. For this reason, this paper makes a step further by identifying some individual characteristics affecting the efficiency of commons. The effects of these factors seem to be coherent with the existing literature.

For instance, the effect of trust on cooperation

-   link of every individual factor with the existing literature

    -   trust: being trusted increase the level of oxytocin, and this impact in turn the cooperative behaviour toward the person who trusted first @Campbell2010 + dispositional trust predicted pro social behaviour + mediated the relation between early-life experience and pro social behaviour @Wu2020 + conditional reciprocity = cooperate just if others do so (strong reciprocity, @Fehr2022, @VanLange1999): if no information on others, we must base our choice in trust

    -   SVO: pro-socials cooperate more because they accord more importance at others' outcome + joint outcome (vs pro-selves = competitors and individualist) @VanLange1999, integrative model of social value orientation (by according importance at the role of equality in outcomes for pro-social, this model allows to explain why pro-social can behave even less cooperatively than pro-self when reciprocity is violated)

    -   @Tajfel1974: classic TIS, explanation of why more group identity = more cooperation, see also experiments with minimal groups paradigm + @Fritsche2018: The collectively interchangeable state of "social identity" makes people susceptible to adopt collective appraisals, goals, and behavior tendencies that are shared within their ingroup as their own

    -   social status: low cooperate more than high =\> theories of Iacoviello etc!!!\
        @Vargas-Salfate2018: more system justification in high-status individuals!

    -   @Janssen2015: confirmation of several factors (to read!!!)

-   group factors: communication linked to social norms and social identity + anonymity underlies all the others + incentives has dark sides

-   saying that these factors are not independent, but interconnected

    -   For example, women level of cooperation is predicted by their trust in their interaction partner, while men's trust and cooperation are not linked [@Irwin2015]. Moreover, women's lower cooperation is explained by the fact that they expected their partner to contribute less to the common good [@Dorrough2019].

    -   @Ackermann2019: link between SVO and trust

    -   @Navarro-Carillo2018: high status more generalized trust, low status more in-group depersonalized trust (because more social networks) =\> link with gender and trust

    -   link between gender and social status, and between social status and social identity

    -   @Aksoy2019: see if class or group had a stronger impact =\> could be a reference to say that it is important to create a common social identity that makes weaker the effect of previous belongings **???**

    -   @Hogg2006: communication and social identity

    -   @Chou2013: link between trust and collective efficacy, so that trust could also fit in SIMPEA and SIMCA

    -   de-individualized trust = more trust in people of in-group just because they are in-group

    -   @VanLange1999 =\> women were more prevalent between pro-socials and less between pro-selves (secondary result)

-   in-group identification is also the main predictor of collective action in SIMCA and SIMPEA

SIMPEA model: [https://www.researchgate.net/profile/Michael-Brenner-Fliesser/publication/368219709/figure/fig2/AS:11431281117565450\@1675469130812/SIMPEA-model-Fritsche-et-al-2018.jpg](https://www.researchgate.net/profile/Michael-Brenner-Fliesser/publication/368219709/figure/fig2/AS:11431281117565450@1675469130812/SIMPEA-model-Fritsche-et-al-2018.jpg)

environmental goals are usually collective goals and can therefore be attained just through a collective action (SIMPEA, @Fritsche2018a) =\> link with the fact that commons are successful only if people cooperate, "These crises, such as climate change, are large-scale and have the structure of common good dilemmas"

in order to evaluate the climate crisis as relevant for the self, individuals need a connection with others who are more likely to be affected by it

4 elements that affect appraisal of environmental crisis + pro-environmental behaviours:

-   in-group identification

-   collective efficacy beliefs =\> trust: since trust are the belief in others' contribution, it must be linked to the perceived collective efficacy (no trust = no collective efficacy, because if one thinks that the others are not contributing as they should then they can not believe that a collective action can be efficient)

-   in-group norms and goals

SIMCA model: [https://www.researchgate.net/profile/Fouad-Bou-Zeineddine/publication/351515886/figure/fig1/AS:1042875005542400\@1625652185052/an-Zomeren-et-al-2008-Social-Identity-Model-of-Collective-Action-SIMCA-with.png](https://www.researchgate.net/profile/Fouad-Bou-Zeineddine/publication/351515886/figure/fig1/AS:1042875005542400@1625652185052/an-Zomeren-et-al-2008-Social-Identity-Model-of-Collective-Action-SIMCA-with.png)

-   individual factors: some of them can be manipulated to improve cooperation, some of them must just be taken into account when "planning"

-   group factors = strategies that groups can implement to improve cooperation

-   =\> talk about the practical implication of this (interventions, ...)

## Limitations, future perspectives and conclusion

This review's aim was to provide an overview of the factors that affect cooperation in social dilemma situations such as commons. However, most of the studies included had a laboratory social dilemma setting, who do not allow to exactly reproduce the dynamics present in a commons [@Cole2010]. Therefore, this papers offers a good basis about elements influencing cooperation in commons, but future studies should confirm this results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.

Another limitation of the current review is the lack of representation of the global population. In fact, two thirds of the analyzed papers were studies conducted in Europe or in North America, and the fourteen reviews were also mostly based on studies from Global North. The current results apply therefore to Western societies and should not be generalized to Global South. In fact, there are cultural differences in the levels of individualism and collectivism: generally, North American and European cultures are more individualistic while South American, African and Asian ones are more collective [@Triandis1989]. These differences affect the extent to which individuals define them-selves as part of a group [@Trafimow1991] and, consequently, their behavior in cooperative situations [e.g. @Kuwabara2014]. For this reason, it is likely that underlying mechanism of cooperation are different for Global South and North: comparative studies between this two regions should be conducted in order to reach a global understanding of cooperation.

Finally, the papers included in this review did not provide a longitudinal perspective on the study of cooperation, leaving therefore some open questions: which factors are more important in the early or in the late stages of cooperation? How the different factors evolve and influence each-others over time? That is why longitudinal investigations about the evolution of commons are needed.

To conclude, this paper offers a first draft of why people do (not) cooperate in commons, and highlights a lack of systematic studies about cooperation in commons. However, it is not without limitations, thus further researches are necessary to reach a deeper understanding of the topic. + say the first important practical implication that we can already deduce (social identity, ...)

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::