### 2023-12-12 scoping_review_short.Rmd
12.12: integration on Vincent's and Lilla's basic feedbacks

---
title             : "Understanding cooperation in commons through social dilemmas' literature: a scoping review"
shorttitle        : "Understanding cooperation in commons"

author: 
  - name          : "Pedrazzini Sabina"
    affiliation   : "1"
    #corresponding : yes    # Define only one corresponding author
    address       : "University of Bern, Centre for Development and Environment (CDE), Mittelstrasse 43, CH-3012 Bern"
    email         : "sabina.pedrazzini@unibe.ch"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
     # - "Conceptualization"
     # - "Writing - Original Draft Preparation"
     # - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Centre for Development and Environment, University of Bern"

#authornote: |

abstract: |
  Commons represent an alternative to the classical capitalist model, therefore they could be the needed change toward a more sustainable society and economy. However, for a commons to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in the scientific field of the commons. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since social dilemma experiments and commons share some structural characteristics, we could gain insights into cooperation in commons and how to implement it thanks to the social dilemmas literature.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, in order to better understand what could shape cooperation in commons. A key-word search and a subsequent selection led to the inclusion of 89 papers, which were used to identify 9 factors that can influence cooperation both at the individual and group level: gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity. These factors are discussed within the existing literature on commons and cooperation. More specifically, the influences that these elements can have on each-other are highlighted, showing that all factors should be taken into account when investigating cooperation in commons. Finally, the practical implication of this work for commons' organization, as well as the review's limitations, are pointed out.

#keywords          : "cooperation, social dilemmas, commons"
#wordcount         : ""

bibliography      : CommonPaths.json

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_word
always_allow_html : true
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

In recent years, the interest in alternative ways of social organization increased drastically due to the social and ecological crises we are facing [@Euler2018]. Commons are one of these alternatives, since they are a way of organizing resources beyond the classic private and public model [@Ostrom1990], and, therefore, they represent an option to the capitalistic system [@Euler2019]. For this reason, commons could be an integral part of the path to attain a more sustainable and fair social organization [@Schmelzer2022].

Commons have been defined in different ways, starting by Elinor Ostrom who, in her seminal work, refers to them as a resource system shared between a self-managed community of users [@Ostrom1990]. This community shares a set of social practices, values and norms. Therefore, it is important to consider the commons not only as a resource, but as the integration of the resource, the community and its social protocols [@Bollier2014]. For this reason, beyond the material aspect, the notion of social practices (i.e. commoning) is essential [@Euler2018]. "Commoning" refers to the process through which people (i.e. the commoners) find solutions for their needs by self-organizing independently from the state or the market, and by creating autonomously managed institutions to govern the shared resource [@Euler2018; @Schmelzer2022]. The community is therefore engaged in defining rules to successfully manage the commons [@Meroni2022]. In conclusion, commons can be understood as "living social systems" that allow commoners to self-organize to address shared problems [@Bollier2019].

These definitions highlight the importance of the community as a part of the commons, and that is why it is essential to have a closer look at the people who form it. More specifically, in order for a commons[^1] to be successful, its members have to cooperate, i.e to incur a personal sacrifice for the greater good [@Kollock1998]. Although it is essential, cooperation can be difficult to achieve due to two characteristics of the commons' social dynamic. First, people can believe that their individual contribution will not have a big impact, and this feeling of personal inefficacy discourages cooperation [@VanVugt1998]. Secondly, in a commons, every member faces a conflict between individual, short-term interests and collective, long-term ones [@Hardin1968]. Act for the common good requires a personal sacrifice, and therefore the temptation to pursue the individual interest is constantly present. For these reasons, we need to understand the factors that affect cooperation, and how we can improve it.

[^1]: The term "commons" is both singular and plural [@Bollier2014]. For the difference between the singular forms "common" and "commons", see @Feinberg2021.

## Cooperating in commons

In the existing literature, there are several explications of why people adopt or not cooperative behaviours in commons. For instance, the Rational Choice Theory predicts that people are too self-interested to cooperate in such contexts. As a consequence, they will inevitably destroy the shared resource, unless an external institution forces them to cooperate [@DeCaro2021; @Hardin1968]. However, there are some less pessimistic theories that predict that people can transcend their personal interest and successfully cooperate in commons. An example is the Greed Efficiency Fairness hypothesis, which posits that the individuals' greed to pursue their narrow interests is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. Similarly, according to the Humanistic Rational Choice Theory, the desire to maintain basic social-psychological needs motivates people to cooperate [@DeCaro2021]. Finally, cooperating in commons can also be a self-presentation strategy [@VanVugt2010a], since showing altruistic behaviours has a positive effect on reputation [@Roberts1998].

On a more practical level, @Ostrom1990 reported several cases of successful commons, where individuals managed to create a system to ensure cooperation. Based on her observations, she established eight design principles that distinguish successful from unsuccessful commons. First, commons need clearly defined boundaries to determine who has access to the shared resource, and the governing rules must take into account the local conditions. Moreover, people affected by these rules must be able to participate in modifying them. A monitoring and a graduated sanctions system must also be implemented, as well as low-cost conflict resolution arena. Finally, external authorities must not challenge the commons' members' right to self-organize[^2] [@Ostrom1990].

[^2]: The last principle is valid only for commons that are part of a larger system, and is about nested enterprises: all the governance activities should be organized in multiple layers of nested enterprises.

There are therefore several studies that focused on commons and on how they can be successful, but social psychological and quantitative perspectives are missing from this literature. In fact, to the authors' knowledge, there are few studies that investigated the causal effect of individual and group factors on people's willingness to cooperate in a real-life commons [e.g. @Noonan2016].

At the same time, an experimental approach is very popular in the laboratory study of social dilemmas, which identified several factors that shape cooperation [e.g. @VanLange2013]. In these experiments, people face a conflict between personal and collective interest: each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. At the same time, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980]. Since many of the current global problems have a social dilemma structure, this topic has been extensively studied in different fields such as psychology, economy, biology, or mathematics [@VanLange2013]. Social dilemmas are very useful to gain insights into how cooperation works, because they can easily be implemented in a laboratory setting, and they provide a standard experimental paradigm to test under which conditions people cooperate [@Balliet2011].

## Current review

Commons and social dilemmas are deeply linked. First, they share the same basic conflict between personal and collective interests [@Hardin1968; @Dawes1980]. For this reason, commons can be considered as a particular case of social dilemma. In fact, @Kopelman2002 used the term of "commons dilemma" to define situations where a non-cooperative behaviour can seriously endanger the future of the shared resource. Moreover, commons have been classified by @Cumming2018 as "take-some dilemmas", a specific kind of social dilemma where people are tempted to increase their personal benefit by taking more than their fair share from a shared resource. Moreover, cooperation and defection in social dilemma laboratory experiments can predict people's behaviour in a real-life commons [e.g. @Reindl2019].

The current paper synthesizes existing research about cooperation in social dilemmas, to provide a set of factors that impact cooperative behaviours in commons. This work will therefore lay the foundations to implement experimental and quantitative studies in real-life commons, which, in turn, will allow to gain insights about cooperation beyond artificial laboratory settings. Other studies already used a social dilemma framework to study real-life commons [e.g. @Hartl2022]. While these studies usually focused on just one particular case, the current review is seeking to extend this approach by using social dilemma literature to gain a global understanding of cooperation in commons.

This scoping review has two main goals. The first is to identify factors on the individual level that can shape cooperation in commons. Moreover, since it has been widely demonstrated that commoners can successfully self-organize to govern a commons [@Ostrom1990; @Larsson2016], the second goal is to understand which strategies groups can implement to encourage their members to cooperate.

# Methods

To conduct this scoping review, we analyzed different scientific publications in response to the research questions. To gather the papers, we performed a Boolean search in Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", and "commons", and each of them was paired with each one of the following: "cooperation", "conflicts", "reciprocity", and "trust". We carried out the research during the month of August, 2023. According to @Munn2022, the type of source was not a selection criterion, therefore journal articles, conference papers and book chapters could be included, to have an overview as complete as possible about the existing literature on the topic. Moreover, we did not use any chronological or geographical limitations, once again to provide a broader vision. However, we implemented some selection criteria concerning the papers' methodology. First, we only selected papers written in English. Moreover, we focused on studies with human participants (or theoretical articles), so we excluded researches using an agent-based model. Secondly, among experimental or correlative studies, we only selected cases with adult participants.

Initially, 5210 papers emerged from the key phrases search. Deleting the duplicates reduced them to 4586. Based on titles and abstracts, we eliminated 4333 articles, leaving 253 texts to examine. After a full-text evaluation of the remaining papers, we selected 89 for the present review. Figure 1 resumes this selection process.

![Figure 1. Overview of the selection process.](images/figure1-01.png)

## Papers' description

The 89 included papers were published between 1984 and 2023, with a median of 2015. Almost all the selected texts were journal articles, with the only exception being a book chapter. Most of them were experimental studies (*N*=60), followed by literature review and meta-analysis (*N*=14) and correlative studies[^3] (*N*=14). The remaining paper was a qualitative study.

[^3]: Studies where at least one variable is manipulated by researchers are part of the category "experimental studies", while "correlative studies" refers to papers using quantitative analysis, but without manipulations.

### Topic

We identified two dimensions in the reviewed papers: setting (lab-based vs real-life) and context (social dilemma vs commons). There were papers in all four combinations of these dimentions, with the most represented category being "lab-based social dilemma" (*N*=75), namely highly controlled laboratory experiments where participants had to take cooperative decision. This was followed by "real-life social dilemma" (*N*=7), which investigated cooperation in different real-life situations with a social dilemma structure such as recycling behaviours [@Rompf2017] or use of public transportation [@VanLange1998]. Finally, "real-life commons" refers to studies of cooperation where participants were members of a commons in their everyday life, while "lab-based commons" refers to studies with a laboratory social dilemma paradigm, but with clear references to a concrete commons [e.g. @Mosler1993].

### Journal

More than a half of papers was published in psychology journals, such as *Group Processes & Integroup Relations* or *Psychological Bulletin* (*N*=34), in economic journals like *Experimental Economics* (*N*=14), or in interdisciplinary ones such as *Journal of Economic Psychology* (*N*=8). The remaining papers where published in journals concerning different topics, such as biology (*N*=6), mathematics and game theory (*N*=4), sustainability (*N*=3), or commons (*N*=2).

### Country

Finally, most of the studies were conducted in Europe (*N*=40) (especially in Germany (*N*=12) and in the Netherlands (*N*=8)), followed by North America (*N*=20, of which *N*=19 in the USA) and Asia (*N*=7, of which *N*=6 in China). Moreover, six studies were set in more than one country, and fourteen were meta-analysis or literature reviews. The two remaining studies took place in Australia and Colombia, respectively.

A resume of the papers' description is provided in Table 1.

# Individual factors

Our analysis identified five individuals factors that impact cooperation: gender, social status, group identification, values and personality traits, and trust. In the following sections, we describe their principal effects on cooperation, and provide an overview of the moderators of each effect. More details about moderators are included in Appendix A.

## Gender

The effect of gender on cooperation is controversial, and studies that investigated it provided contradictory results. Sometimes men cooperated more [@Dorrough2019], while sometimes women did [@Peshkovskaya2017]. However, two meta-analyses reported a non significant effect size [@Balliet2011; @Spadaro2023]. This indicates that gender differences in cooperation are very flexible and depend on the context [@Spadaro2023]. More specifically, @Sell2009 explained these contradictory results "by structural differences [of the experimental settings] and identities triggered by those differences" (p. 322).

The effect of gender on cooperation is therefore subjected to several moderators, since the situation in which women and men interact is essential to determine which gender will cooperate more. For instance, men are more cooperative when they know other participants [@Peshkovskaya2017], when they feel part of a team [@Gomez-Ruiz2020] or when they interact repeatedly with the same group [@Balliet2011]. The partners' gender is also an important moderator. Men-only groups are more cooperative than women-only [@Colman2018; @Peshkovskaya2019], while women cooperate more in mixed groups [@Balliet2011; @Peshkovskaya2019]. In general, groups with a majority of women display the highest levels of cooperation [@Spadaro2023], even if this effect of group composition can be moderated by other factors [e.g. @Barrero-Amortegui2021]. Moreover, men are more sensitive to greed incentives than women , and thus they cooperate less in situations with strong temptations to free-ride are present [@Simpson2003]. On the other hand, women's principal motive in cooperative interaction is avoiding to be exploited, and that is why they cooperate less if they don't trust other participants or if they believe than others provide small contributions [@Irwin2015; @Dorrough2019].

Lastly, gender differences in cooperation could be partly explained by social status, where women represent the low-status group [@Sell2009]. For this reason, it is important to gain insights into how social status impacts cooperation.

## Social status

The results of studies investigating the impact of social status on cooperation do not allow us to draw a clear conclusion about this effect. For instance, some findings suggest that low-status participants cooperate more than high-status ones by contributing a higher proportion of their endowments [@Kingsley2016]. However, @VanLange2013 identified some studies where wealthier people cooperated more.

Social status can also create inter-class conflicts, because people tend to cooperate more with the members of the same social class than with members of other classes [@Aksoy2019]. This is especially true for low-status groups, where solidarity can arise either in the form of discrimination toward high-status members [@Camera2020], or in the form of fewer punishments for other low-status [@Chen2022].

Regarding the effect of overall group composition, studies disagree. On one hand, @Aksoy2019 found that groups with an inequality of status between members displayed lower levels of cooperation than homogeneous groups. On the other hand, economic and socio-cultural heterogeneity are only negatively correlated with in-group trust, while there is no evidence of their effect on the integrity of a common-pool resource [@VanKlingeren2021].

Taken together, these results suggest that the impact of social status on cooperation is complex. Moreover, similarly to what @Sell2009 reported about the effects of gender and social status, @Aksoy2019 demonstrated that the lower level of cooperation between members of different social classes is mainly due to actors' identification with their classes. This shows that social identity and social status are deeply connected, and that it is difficult to separate their effects on cooperation.

## Group identification

Social identity has a strong, positive effect on cooperation: the more individuals identify with a group, the more they cooperate with other members of the group [@VanVugt2009], for instance by providing more voluntary activities for the collective good [@Noonan2016]. In fact, humans have a very developed sense of belonging, and when they strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they value more group-related goal [@Arora2016]. A strong identification also maintains high levels of cooperation over time [@Zhang2019]. Finally, further confirmations of the strength of this effect are the fact that it is the same for individualist and collectivist people [@Chen2007] and that group identification can be easily implemented in a laboratory experiment and still have a strong effect on cooperation [@Wit1992].

In general, people cooperate more with other in-group than with out-group members [@Dorrough2015; @Aksoy2019], but the individual differences in the strength of identification can lead to inequalities in cooperation. In fact, individuals who strongly identify with their group can even compensate, by cooperating more, for other in-group members' defection [@VanVugt2009; @Arora2016].

Intragroup cooperation can be enhanced via a competition between different groups, because in such a context individuals' and groups' interests are aligned [@Puurtinen2009]. Moreover, a study showed that intergroup competition increased cooperation within the group, while individual, dispositional competitiveness had the opposite effect [@Nockur2020]. This demonstrated that the increase of cooperation via competition is due to group membership factors, and not to individuals' competitiveness.

Despite the beneficial effects of inducing intergroup competition, this could be harmful in real-life situations, where resources are shared between several groups. In fact, there is a greater risk of over-exploitation when an intergroup competition is present [@VanVugt2009]. To maintain the positive effect of group identification while eliminating the risks associated with competition, it is useful to prime a super-ordinate identity [@VanVugt2009]. This can be done by making salient the common characteristics of the different groups rather than their differences [@Kramer1984] or by inducing the feeling of a common fate [@Zhang2019].

Finally, social identification can also moderate other factors' impact on the willingness to cooperate, such as monetary incentives [@VanVugt2009] or a previous group failure [@Jackson2011; @Jackson2012].

## Values and personality traits

Personal values and personality traits play an important role in predicting cooperative behaviour [@Hilbig2018]. One of the most studied is the Social Value Orientation (SVO), that divides people in pro-social and pro-self based on how much they value their own and others' outcomes in situations of interdependence [@Balliet2009]. The effect of SVO on cooperation has been widely demonstrated: in a meta-analysis, 81 out of the 82 papers examined reported that pro-social cooperated significantly more than pro-self [@Balliet2009].

There are several moderators to this main effect. For instance, when economic incentives to cooperation are present, pro-self individuals attain a level of cooperation similar to pro-social's one [@Emonds2011; @Balliet2009]. Similarly, in a study about commuting habits, the extent to which the cooperative behaviour (i.e. choosing to commute by public transportation or carpooling) was perceived as efficient impacted pro-self individual's decision more than pro-social's [@VanLange1998]. Another moderator is the uncertainity about the group's size: if the uncerainty is high pro-social, but not pro-self, increse their level of cooperation [@DeKwaadsteniet2008]. Finally, pro-self increase their cooperation if they have the possibility to deliberate about the situation [@Lu2019].

Other studies also focused on how much people are self- or social-orientation impact cooperation by using other constructs. More cooperative individuals have an interdependent definition of the self [@Liu2009a] and see themselves as part of the group [@Probst1999]. Moreover, allocentric (i.e. collectivist) people are more sensitive to cooperative social norms [@Chen2007]. Finally, other personal values and traits that affect cooperation include pro-environmental values [@Sussman2016], political orientation [@Fosgaard2019], humility and honesty [@Kieslich2014; @Hilbig2018], self-control [@Kocher2017], and justice sensitivity [@Schlosser2018].

## Trust and beliefs about others' contribution

Trust can be defined as the "expectations of others' benevolent motives in situations that involve a conflict between self and collective interests" [@Balliet2013, p.1]. There are two categories of trust: first, dispositional trust is an individual general trait that can be assessed with a personality scale. Secondly, state trust is linked to the context and represent the expectations about a partner's behaviour in a specific situation [@Balliet2013]. The two are positively correlated [@Lubke2021], and both have a positive effect on cooperation [@VanLange2013; @Balliet2013].

Several studies have confirmed this relationship. For instance, high levels of trust are associated with more contributions to a public good [@Kocher2015], individuals' cooperation can be predicted by their beliefs about others' contributions [@Smith2013], people who are pessimistic about others' contribution cooperate significantly less than optimistic people [@Oyediran2018], and individuals who cooperate more display higher levels of trust in following interactions [@Chaudhuri2002]. The enhancing effect of trust on cooperation has also been demonstrated in real-life studies [see for example @Sturm2019; @Franzen2019]. Finally, trust matters more in situations involving a large, rather than small, conflict between personal and collective interests [@Balliet2013], and when people are uncertain about others' intentions [@VanLange2013].

People can also display different levels of trust in the "reliability, effectiveness and legitimacy of public institutions" [@Rompf2017, p. 2]. Institutional trust is associated with the probability of recycling, because knowing that sanctions can be efficiently applied motivates people to comply with norms.

# Organizational factors

On the organizational level, we identified four factors from the reviewed papers: incentives, communication, social norms and anonymity. As for the individual factors, in the following sections we detail the main effects of each factor, and we provide an overview of possible moderators. More details can be found in Appendix A.

## Incentives

Cooperation can be modified by introducing incentive systems, that can be either punishments or rewards. Both have a positive effect on cooperation [@Balliet2011a]. However, some studies found that punishments are more efficient than rewards [e.g. @Milinski], while others came to the opposite conclusion [e.g. @Chen2022].

Several factors moderate the main effect of incentives on cooperation. A meta-analysis found that both rewards and punishments are more effective if their implementation is costly rather than free, and during iterated, rather than single, interactions [@Balliet2011a]. Moreover, punishments work better in groups where all members have the same income [@Kingsley2016], because in heterogeneous groups a form of solidarity between low-status members arises and encourages free-riding [@Chen2022]. Punishments are also more effective if participants are friends [@Balliet2011a] and when only cooperators have the power to punish defectors [@Xiao2016]. Finally, the link between behaviour and outcome must be clear [@Xiao2016], and the punishment must have a significant consequence [@Egas2008].

In general, there are more studies on punishments than on rewards [@VanDijk2015]. When a punishment system is present, participants' perceive the situation as more fair [@Nockur2021]. However, punishments must be applied with carefulness. Ideally, groups should not punish in the early stages of the interaction, and have a norm of when and how to punish [@Rockenbach2019]. Moreover, a punishment system could entail a second-order social dilemma, since cooperators are better-off by not investing resources into punishing, if all others do [@Milinski2012]. Also, if there are no rules about who can punish, non-cooperators can punish cooperators as a form of retaliation [@Xiao2016; @VanDijk2015]. A democratic punishment system, where a fine is given only if a majority agree, could solve this problem [@Pfattheicher2018].

Despite enhancing cooperation, incentives have negative sides: they reduce participants' intrinsic motivation [@VanDijk2015; @Schlosser2018] by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. They also undermine participants' trust in other members' internal motivation to cooperate. Therefore, if they are removed, cooperation drops [@Irwin2014; @Mulder2006].

Finally, incentives are often studied in economic formats, but they can also be social. In fact, expecting honour or shame (after cooperating or defecting, respectively), or expressing disapproval toward defectors can enhance cooperation [@Jacquet2011; @Nelissen2013]. An ostracism system where participants can vote to exclude other members is also very effective [@Maier-Rigaud2010]. In conclusion, social sanctions can be even more efficient than economic ones, because cooperation do not drop after their removal [@Nelissen2013].

## Communication

Communication between members has a general positive effect on cooperation [e.g. @Balliet2010], because it allows to shape a cooperative social norms [@Janssen2014; @Bicchieri2002] and to reach a consensus about the best strategy for the group [@Hopthrow2005]. In fact, when communication is used to foster a common social identity, promote cooperative social norms and ensure that every member publicly commits, cooperation is higher [@Adams2022].

Concerning the moderation effects, oral communication is more powerful than written messages, and communication has a stronger impact in large, rather than small, groups [@Balliet2010]. Moreover, if communicating requires a personal cost, its effect is weaker than if communication is free [@Janssen2014]. The content of the message is also essential: to be effective, communication must allow participants to become aware of the problem faced by the group, identify different strategies and agree on the one to use. In addition, casual chit-chat also plays a very important role in increasing cooperation [@Koessler2021]. Finally, communication before or during the cooperative interaction have both a positive impact on cooperation [@Balliet2010], while knowing that it will be possible to communicate after the interaction does not enhance cooperation [@Kumakawa2013], and can even have a negative effect [@Torsvik2011].

A particular form of communication is gossip, i.e. the spread of reputational information about another person [@Feinberg2014]. Gossip increases cooperation by providing information about the past behaviour of potential partners, which helps to select the most cooperative members in the following interaction. Moreover, knowing that gossip is possible motivates people to cooperate more to avoid being the target of such gossip [@Feinberg2014]. Gossip can also have the pro-social function of protecting others from potential exploitation [@Feinberg2012]. Finally, providing correct information while gossiping helps to build a positive reputation [@Giardini2021] and reduces negative affects on the gossip actor [@Feinberg2012].

## Social Norms

Social norms can provide a guide on how to behave, especially when there is uncertainty about others' behaviour like in cooperative situations [@VonBorgstede2018]. Both injunctive (what others approve) and descriptive (how others actually behave) norms impact cooperative behaviours [@Thogersen2008]. Moreover, social norms help to maintain cooperation, because individuals who do not respect them are susceptible to social sanctions [@Rege2004].

Social norms shape cooperation because, when taking a decision, people ask themselves: "What does a person like me do in a situation like this?" [@Arora2012]. If they define the situation as a cooperative task and strongly identify with the group, group's norms will be more accessible and they will have a stronger impact on behaviour [@VanLange2013]. Moreover, people have a general tendency to imitate what others do, even if others' actions do not have an impact on their personal gains [@Bardsley2005a].

Finally, social norms can be modified by participants' behaviour. For instance, the presence of consistent cooperators (i.e. people who always cooperate, independently from others' actions) creates a more cooperative social norm and thus increases cooperation [@Weber2008]. Moreover, participants can also shape social norms via communication [@Bicchieri2002].

## Anonymity

In general, in non-anonymous situations cooperation is significantly higher than in anonymous ones [@Wang2017]. In such situations there are more social incentives to cooperate, because people worry about others' approval [@Rege2004]. Moreover, several factors that can increase cooperation rely on a reduction of anonymity [@Wang2017]: for instance, punishing defectors is impossible without knowing who they are. When behaviours are visible, cooperation can also be a self-presentation strategy to be perceived as pro-social and therefore to increase one's social status. In fact, the positive effect of non-anonymity is stronger in situations where cooperation bears a cost without offering personal gain [@Butz2021; @VanVugt2010a].

Therefore, reducing anonymity in cooperative interaction could significantly improve cooperation. A possible way to do this is to reduce the group's size, because larger groups lead to de-individuation and it is thus more difficult to identify defectors [@Romano2016]. In fact, cooperation levels are higher in small, compared to large, groups [@Jiang2021]. Anonymity can also be reduced by making people publicly promise to cooperate. Such a commitment significantly increases cooperation rates, even if after committing people are not forced to cooperate [@Mosler1993; @Przepiorka2020]. Even a compulsory commitment has a beneficial effect on cooperation [@Koessler2021a].

# Discussion

The aim of the present paper was to synthesize the existing literature about cooperation in social dilemmas, in order to identify factors that could impact cooperative behaviours in commons. Through a scoping review approach, we identified nine factors, both at the individual and group level. Our results seem to be in line with others findings about cooperation in commons. For instance, factors such as gender, social status, SVO, incentives, and communication were also found by @Kopelman2002 in their review about lab-based commons, and their effects were the same as the ones reported in the current paper.

Moreover, some of the factors can be related to Ostrom's principles to successfully manage a common resource [@Ostrom1990]. For instance, the clearly defined boundaries can strengthen commoners' social identification, since only when an in- and an out-group exist social identity is salient and therefore an identification is possible [@Tajfel1974]. Even if Ostrom developed this principle on a more organizational level, its secondary effect of creating a social identity can also contribute to improve cooperation. Secondly, communication is essential for the principle about the creation and modification of norms [@Bicchieri2002], and for the one about the arenas to solve conflicts. Finally, the principles concerning the monitoring and the graduated sanction systems underline the importance of incentives. The monitoring system refers to social incentives, and is linked to anonymity, since if people know that their behaviour can be observed, they are more willing to conform to social norms because they worry about others' approval [@Nelissen2013; @Rege2004]. On the other hand, a sanction system can also have a positive impact on cooperation [@Balliet2011a], especially if it is graduate, so if sanctions are not too high at the first violation [@Rockenbach2019].[^4]

[^4]: Principles 2, 7 and 8 refer to elements that are not pertinent to the current review, and therefore do not match any of the factors identified.

Ostrom's principles are especially related to factors at the organizational level. This is understandable when considering her economic approach to the study of the commons. However, taking into account the social dynamics and the characteristics of the commoners is also essential to have a better understanding of how these communities work [@Euler2018; @Nightingale2011]. For this reason, this paper went a step further and identified five individual characteristics affecting the efficiency of commons. Although less studied in commons literature, these individual factors are consistent with existing theories.

First, the reported positive effects of trust on cooperation is explained through the principle of reciprocity, that posits that individuals cooperate more facing cooperative partners [@Fehr2002; @VanLange1999]. Since in real-life situations it is not always possible to know exactly others' willingness to contribute, trust (i.e. beliefs about how much others cooperate) becomes essential to decide how to behave [@Vuolevi2012]. Moreover, trust among participants of a common-pool setting is at core of the framework of @Poteete2010, aimed at explaining why commons can be successful.

Secondly, the influence of SVO on cooperation is due to the difference between pro-social and pro-self in the importance accorded to the relative benefits. More specifically, pro-social value more others' outcome and equality in outcomes than pro-self [@VanLange1999]. Moreover, the difference in weights accorded to the personal and others' outcome seem to be the only dissimilarity between pro-social and pro-self, because they process the cooperative interaction in the same way [@Fiedler2013].

The reviewed results did not provide a clear conclusion about gender effect on cooperation. However, existing theories would predict that women cooperate more than men, because, according to gender stereotypes, women are more communal and caring, while men are more agentic and performance-oriented [@Eagly2012]. In fact, groups partly composed by women display more solidarity and collaboration than all-men groups [@Westermann2005]. However, every experimental paradigm can trigger the participants' gender identity differently [@Sell2009], and for this reason this effect is subjected to several moderators.

Concerning social status, the analyzed studies showed that usually members of low-status groups cooperated more and showed more solidarity with others low-status members than high-status. Although this effect has not been extensively studied, these results seem to be consistent with the existing literature on social status. First, there are structural differences between high- and low-status group's dynamic, that make high-status members more individualistic than low-status [@Lorenzi-Cioldi2002]. High-status individuals also display more system justification and more Social Dominance Orientation than low-status [@Caricati2012; @Vargas-Salfate2018]: they see the social system as fair and they have a preference for inequality between social groups. This is negatively correlated with concern about others and altruism [@Pratto1994].

The last individual factor was group identification, and the reviewed results showed that the more individuals identify with their group, the more they are likely to act for the group's good. In fact, according to the Social Identity Theory, people's identity is also composed by their group affiliations, and the stronger they identify with a social group, the more important this group becomes for their identity and the more their personal interests aligns with group's ones [@Tajfel1974]. Also, within a social group, individuals perceive themselves as more similar to other members, and therefore they are more susceptible to adopt collective goals and behaviours [@Turner1984; @Fritsche2018]. Moreover, strongly identified people are less likely to leave a group in case of failure or difficulties [@Ellemers1997]. Thus, a strong identification is essential for a group's success. In-group identification is also essential to engage in collective actions [@VanZomeren2008], and it is the main predictor of pro-environmental actions in the Social Identity Model of Pro-Environmental Action [SIMPEA, @Fritsche2018].

## Factors' interactions

Until now, the identified factors have been discussed separately. However, their impacts are interconnected, as already anticipated in some reviewed results about moderation effects. For instance, there are gender differences in trust and SVO: trust is a better predictor of cooperation in women than in men [@Irwin2015; @Dorrough2019], and women are more prevalent between pro-social and less between pro-self [@VanLange1999]. Moreover, SVO and trust are linked [@Ackermann2019], and the expectancy about others' cooperation can even be seen as a part of the construct of SVO [@Bogaert2008]. Therefore, pro-self individuals show less cooperation in part because they expect others not to cooperate.

Trust also depends on social status: high-status individuals display more generalized trust (i.e. general beliefs on others' benevolence), while low-status more depersonalized trust (i.e. more trust in other members of their in-group, just because of their membership) [@Navarro-Carrillo2018]. Since high-status individuals perceive the system as fair [@Caricati2012], it is easier for them to believe in others' benevolent intentions. Conversely, belonging to the group is more important for low-status than for high-status [@Lorenzi-Cioldi2002], therefore they see other group's members as more trustworthy, even without knowing them personally.

Secondly, communication impacts several other factors. For instance, it plays an essential role in shaping social norms [@Bicchieri2002] and it can improve cooperation by signalling a person's intentions and trustworthiness [@Janssen2015]. Moreover, the opportunity to exchange about the strategies that need to be implemented improve people's commitment within a group, that is a component of group identification [@Postmes2001]. At the same time, group identification also moderates other effects. First, as a consequence of the Social Identity Theory, high-identified individuals are more willing to comply to the group's norms and to pursue the group's goal [@Tajfel1974]. Therefore, they need less cooperation-enhancing measures such as incentives [@VanVugt2009] or non-anonymous conditions [@Barreto2000]. Moreover, social rewards work better with people who strongly identify with the group that can approve their behaviour [@Kopelman2002].

Finally, as mentioned in the results, anonymity underlies several other factors. First, people comply more to social norms when actions are accountable, because their concern about others' approval [@Nelissen2013; @Rege2004]. Also, in large groups [i.e. in more anonymous situations, @Romano2016] it is more difficult to asses the trustworthiness of other participants than in small ones [@Janssen2015].

## Theoretical implications

The analyzed papers showed that people often cooperate. This contradicts the Rational Choice Theory (RCT), who posits that individuals are too self interested to sacrifice their personal benefit to the greater good, unless if forced to do so. Our general results are therefore consistent with the existing literature, since RCT has already been refuted by experimental studies, who demonstrated that cooperation was higher than predicted in social dilemmas, and that it was impacted by variables that should not play a role according to this theory [@Ostrom1998].

Moreover, this review also corroborated the lack of quantitative studies about cooperation in real-life commons: of the 89 analyzed papers, just three had these characteristics, and they all were correlative. Therefore, they could not draw a causal link between the investigated factors and cooperative behaviours. This confirms a gap in the literature, that could be filled by futures experimental studies about cooperation in commons. Moreover, this review laid the foundations to implement such experiment in real-life commons, by providing a set of factors that could be tested.

## Practical implications

This review can be used as a guide for commons members to design campaigns or activities to improve cooperation. In fact, some of the identified factors can be manipulated in order to create better conditions for cooperation to arise. Moreover, since the different elements are linked, they can impact cooperation both directly and indirectly, via mediation or moderation effects. Based on the quantity of connections with other elements, the factors that seem to have a stronger effect are trust, communication, group identification and accountability of actions. Therefore, interventions based on this elements can be very powerful tools to improve cooperation in commons. On the other hand, some individual factor cannot be manipulated, but must be taken into account when designing an intervention. For instance, commons' composition in terms of gender and social status can have a large influence on the overall cooperation.

# Limitations and future perspectives

To conclude, this review summarized existing researches about cooperation in social dilemmas, and therefore provided a set of factors that can help better understanding cooperative behaviours in commons. It also highlighted a lack of quantitative studies about this topic in real-life commons. It is, however, not without limitations.

First, there is a lack of representation of the global population. In fact, two thirds of the analyzed papers were studies conducted in Europe or in North America, and the fourteen reviews included were also mostly based on studies from Global North. The current results apply therefore to Western societies and should not be generalized to Global South. In fact, there are cultural differences in individualism and collectivism: generally, North American and European cultures are more individualistic, while South American, African and Asian ones are more collective [@Triandis1989]. These differences affect the extent to which individuals define themselves as part of a group [@Trafimow1991] and, consequently, their behavior in cooperative situations [e.g. @Kuwabara2014]. It is therefore likely that cooperation relies on different mechanisms in Global South and North: comparative studies between this two regions should be conducted in order to reach a global understanding of cooperation.

Moreover, our aim was to use results of social dilemma researches to gain a better understanding of commons, therefore most of the included studies had a laboratory setting. In fact, this kind of setting is very useful to study social dilemmas because it provides a standard paradigm [@Balliet2011]. However, this does not allow to exactly reproduce the dynamics of a commons [@Henry2011; @Cole2010]: some elements such as cultural factors and specific ecological conditions affecting the decision-making process cannot be reproduced in lab experiments [@Janssen2015]. Therefore, this paper offers a good basis about elements influencing cooperation in commons, but future studies should confirm these results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.

Finally, the papers included in this review did not provide a longitudinal perspective on the study of cooperation, leaving therefore some open questions: which factors are more important in the early or in the late stages of cooperation? How the different factors evolve and influence each-others over time? To answer these questions, longitudinal investigations are needed.

In conclusion, a successful commons relies on the cooperation of its members, but this topic has not yet been extensively studied. Thus, further researches are needed to address the limitations of the current review and to reach a deeper understanding of this topic.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::